On the use of 'boilerplate' statistical analysis sections?

### 2021-05-13 manuscript/asa_template/asa_template.Rmd
update manuscript methods results

<!-- For records that did not pass this second stage filter, we reviewed where keywords appeared in the full-text to ascertain the extent of missing statistical analysis sections. ...[TODO: finish this whilst draft is with coauthors]-->

@@

_Details to come_

We applied the clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50. Clustering solutions were assessed using silhouette score, within-cluster dispersion and between-cluster dispersion.

For each dataset, records assigned to invididual clusters were examined for evidence of boilerplate text in two ways. We first reviewed the strongest matches to each cluster based on topic scores[may not be the right word]. We also compared similarities in text between pairs of records assigned to each cluster. Similarities were determined from calculating cosine similarity scores, with comparisons limited to the top 500 records per cluster. Higher similarity score represent greater overlap in text content between statistical methods sections. Results were transformed to lower case for the clustering, but examples are given using the original capitalisation.

>>>>

@@

Let $P \in R^{M \times N}$ denotes the paper content matrix where the  statistical methods sections of $N$ papers consist of $M$ distinct terms. Text clustering for identifying common writing themes in these papers requires this matrix to represent with a vector space model. Let Matrix $P$ be modeled with the unique terms in the processed sections represented with the tf\*idf (term frequent * inverse document frequency) weighting schema to consider both common and rare terms.

Text clustering is known to face the curse of dimensionality due to the high number of terms in doc$\times$term matrix representation [@aggarwal2012mining, @sutanto2018fine]. Therefore text-based methods based on distance, density or probability face difficulties [@park2018examining,@mohotti2018corpus,@mohotti2019concept]. Specifically, the distance difference between near and far points becomes negligible in high-dimensional data [@aggarwal2012mining]. This directly affects the distance-based methods such as \textit{k}-means \cite{jain} in accurately identifying the common subgroups. In addition, the sparseness of this high dimensional matrix representation does not allow to differentiate the user groups based on density differences [@mohotti2018corpus, @mohotti2018efficient].

Non-negative Matrix factorization (NMF) which maps the high-dimensional data to a lower-dimensional space has been found to provide an effective solution by allowing to form clusters in the lower-dimensional space [@kim2014algorithms,@luong2019clustering]. NMF approximates topic vectors in linear manner considering the context of terms.

In traditional NMF [@aggarwal2012mining], the high-dimensional matrix \( P \in R^{M \times N} \) is approximated by learning two factor matrices \( W \in R^{M \times g} \) and \( H \in R ^{N \times g} \)  where $g$ is the number of cluster groups or common topics in the data collection, as follows.

\begin{equation}
\label{eq:2}
 P \approx WH^{T}
\end{equation}

The matrix factorization process approximates the lower dimensional non-negative factor matrices $W$ and $H$ such that they can represent high dimensional $P$ with the least error. The objective function in Eq \ref{eq:Eq2} iteratively attempts to update the entries in $W$ and $H$ to find the optimum values which possess minimum sum of square error for all the elements in both of those matrices. NMF uses Frobenius norm as its objective function and will find the optimum value of \textit{W} and \textit{H} iteratively.

\begin{equation}
\label{eq:Eq2}
\min \frac{1}{2}\|P - WH\|= \sum _{i=1}^{M}\sum _{j=1}^{N} \left(  P_{i,j} -\left(WH \right)_{i,j} \right)^{2}
\end{equation}

Matrix \textit{H} contains the information regarding topic membership of each document. Topic membership of each method section in $P$ is obtained considering the maximum coefficient value in $H$ for a method section.

We applied the NMF clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50. Clustering solutions were assessed using silhouette score, within-cluster dispersion and between-cluster
dispersion.

\subsection{Within topic analysis}

Results from the clustering algorithm were visualised by word clouds, which summarised frequently occurring terms associated with topic membership. For both _PLOS ONE_ and ANZCTR datasets, we reviewed the strongest matches to each cluster based on topic coherence.

Within the _PLOS ONE_ dataset, we used the strongest matches to investigate evidence of boilerplate text, by comparing text with other studies assigned to the same topic. Comparisons were carried out by calculating Jaccard similarity scores between studies at the sentence level. We chose the Jaccard score as an intuitive measure, which summarises the similarity between two sentences by the number of words common to each sentence (intersection), divided by the number of words that appeared in either the strongest match or the sentence being compared. Boilerplate text was defined based on word count (plus or minus 3 words) and a Jaccard score of 0.9 or higher.

We also compared similarities in text between pairs of records assigned to each cluster. Similarities were determined from calculating cosine similarity scores, with comparisons limited to the top 500 records per cluster. Higher similarity score represent greater overlap in text content between statistical methods sections. Results were transformed to lower case for the clustering, but examples are given using the original capitalisation.

### 2021-05-13 manuscript/asa_template/asa_template.tex
update manuscript methods results

\emph{Details to come}

We applied the clustering algorithm to the cleaned dataset, varying the
number of clusters from 1 to 50. Clustering solutions were assessed
using silhouette score, within-cluster dispersion and between-cluster
dispersion.

For each dataset, records assigned to invididual clusters were examined
for evidence of boilerplate text in two ways. We first reviewed the
strongest matches to each cluster based on topic scores{[}may not be the
right word{]}. We also compared similarities in text between pairs of
records assigned to each cluster. Similarities were determined from
calculating cosine similarity scores, with comparisons limited to the
top 500 records per cluster. Higher similarity score represent greater
overlap in text content between statistical methods sections. Results
were transformed to lower case for the clustering, but examples are
given using the original capitalisation.

>>>>

Let \(P \in R^{M \times N}\) denotes the paper content matrix where the
statistical methods sections of \(N\) papers consist of \(M\) distinct
terms. Text clustering for identifying common writing themes in these
papers requires this matrix to represent with a vector space model. Let
Matrix \(P\) be modeled with the unique terms in the processed sections
represented with the tf*idf (term frequent * inverse document frequency)
weighting schema to consider both common and rare terms.

Text clustering is known to face the curse of dimensionality due to the
high number of terms in doc\(\times\)term matrix representation
\citep[\citet{sutanto2018fine}]{aggarwal2012mining}. Therefore
text-based methods based on distance, density or probability face
difficulties
\citep[\citet{mohotti2018corpus},\citet{mohotti2019concept}]{park2018examining}.
Specifically, the distance difference between near and far points
becomes negligible in high-dimensional data \citep{aggarwal2012mining}.
This directly affects the distance-based methods such as
\textit{k}-means \cite{jain} in accurately identifying the common
subgroups. In addition, the sparseness of this high dimensional matrix
representation does not allow to differentiate the user groups based on
density differences
\citep[\citet{mohotti2018efficient}]{mohotti2018corpus}.

Non-negative Matrix factorization (NMF) which maps the high-dimensional
data to a lower-dimensional space has been found to provide an effective
solution by allowing to form clusters in the lower-dimensional space
\citep[\citet{luong2019clustering}]{kim2014algorithms}. NMF approximates
topic vectors in linear manner considering the context of terms.

In traditional NMF \citep{aggarwal2012mining}, the high-dimensional
matrix \(P \in R^{M \times N}\) is approximated by learning two factor
matrices \(W \in R^{M \times g}\) and \(H \in R ^{N \times g}\) where
\(g\) is the number of cluster groups or common topics in the data
collection, as follows.

\begin{equation}
\label{eq:2}
 P \approx WH^{T}
\end{equation}

The matrix factorization process approximates the lower dimensional
non-negative factor matrices \(W\) and \(H\) such that they can
represent high dimensional \(P\) with the least error. The objective
function in Eq \ref{eq:Eq2} iteratively attempts to update the entries
in \(W\) and \(H\) to find the optimum values which possess minimum sum
of square error for all the elements in both of those matrices. NMF uses
Frobenius norm as its objective function and will find the optimum value
of \textit{W} and \textit{H} iteratively.

\begin{equation}
\label{eq:Eq2}
\min \frac{1}{2}\|P - WH\|= \sum _{i=1}^{M}\sum _{j=1}^{N} \left(  P_{i,j} -\left(WH \right)_{i,j} \right)^{2}
\end{equation}

Matrix \textit{H} contains the information regarding topic membership of
each document. Topic membership of each method section in \(P\) is
obtained considering the maximum coefficient value in \(H\) for a method
section.

We applied the NMF clustering algorithm to the cleaned dataset, varying
the number of clusters from 1 to 50. Clustering solutions were assessed
using silhouette score, within-cluster dispersion and between-cluster
dispersion.

\subsection{Within topic analysis}

Results from the clustering algorithm were visualised by word clouds,
which summarised frequently occurring terms associated with topic
membership. For both \emph{PLOS ONE} and ANZCTR datasets, we reviewed
the strongest matches to each cluster based on topic coherence.

Within the \emph{PLOS ONE} dataset, we used the strongest matches to
investigate evidence of boilerplate text, by comparing text with other
studies assigned to the same topic. Comparisons were carried out by
calculating Jaccard similarity scores between studies at the sentence
level. We chose the Jaccard score as an intuitive measure, which
summarises the similarity between two sentences by the number of words
common to each sentence (intersection), divided by the number of words
that appeared in either the strongest match or the sentence being
compared. Boilerplate text was defined based on word count (plus or
minus 3 words) and a Jaccard score of 0.9 or higher.

We also compared similarities in text between pairs of records assigned
to each cluster. Similarities were determined from calculating cosine
similarity scores, with comparisons limited to the top 500 records per
cluster. Higher similarity score represent greater overlap in text
content between statistical methods sections. Results were transformed
to lower case for the clustering, but examples are given using the
original capitalisation.

### 2021-05-13 manuscript/draft_manuscript.Rmd
update manuscript methods results

_Details to come_

Results were transformed to lower case for the clustering, but examples are given using the original capitalisation.

@@

```{r,fig.width=10,fig.cap="\\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters"}

include_graphics("figures/plosone.10topics.jpg")

>>>>

\subsection{Clustering algorithm}

Let $P \in R^{M \times N}$ denotes the paper content matrix where the  statistical methods sections of $N$ papers consist of $M$ distinct terms. Text clustering for identifying common writing themes in these papers requires this matrix to represent with a vector space model. Let Matrix $P$ be modeled with the unique terms in the processed sections represented with the tf\*idf (term frequent * inverse document frequency) weighting schema to consider both common and rare terms.

Text clustering is known to face the curse of dimensionality due to the high number of terms in doc$\times$term matrix representation [@aggarwal2012mining, @sutanto2018fine]. Therefore text-based methods based on distance, density or probability face difficulties [@park2018examining,@mohotti2018corpus,@mohotti2019concept]. Specifically, the distance difference between near and far points becomes negligible in high-dimensional data [@aggarwal2012mining]. This directly affects the distance-based methods such as \textit{k}-means \cite{jain} in accurately identifying the common subgroups. In addition, the sparseness of this high dimensional matrix representation does not allow to differentiate the user groups based on density differences [@mohotti2018corpus, @mohotti2018efficient].

Non-negative Matrix factorization (NMF) which maps the high-dimensional data to a lower-dimensional space has been found to provide an effective solution by allowing to form clusters in the lower-dimensional space [@kim2014algorithms,@luong2019clustering]. NMF approximates topic vectors in linear manner considering the context of terms.

In traditional NMF [@aggarwal2012mining], the high-dimensional matrix \( P \in R^{M \times N} \) is approximated by learning two factor matrices \( W \in R^{M \times g} \) and \( H \in R ^{N \times g} \)  where $g$ is the number of cluster groups or common topics in the data collection, as follows.

\begin{equation}
\label{eq:2}
 P \approx WH^{T} 
\end{equation}

The matrix factorization process approximates the lower dimensional non-negative factor matrices $W$ and $H$ such that they can represent high dimensional $P$ with the least error. The objective function in Eq \ref{eq:Eq2} iteratively attempts to update the entries in $W$ and $H$ to find the optimum values which possess minimum sum of square error for all the elements in both of those matrices. NMF uses Frobenius norm as its objective function and will find the optimum value of \textit{W} and \textit{H} iteratively.

\begin{equation}
\label{eq:Eq2}
\min \frac{1}{2}\|P - WH\|= \sum _{i=1}^{M}\sum _{j=1}^{N} \left(  P_{i,j} -\left(WH \right)_{i,j} \right)^{2}
\end{equation}

Matrix \textit{H} contains the information regarding topic membership of each document. Topic membership of each method section in $P$ is obtained considering the maximum coefficient value in $H$ for a method section.

We applied the NMF clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50. Clustering solutions were assessed using silhouette score, within-cluster dispersion and between-cluster
dispersion.

For each dataset, records assigned to invididual clusters were examined for evidence of boilerplate text in two ways. We first reviewed the strongest matches to each cluster based on topic coherence. We also compared similarities in text between pairs of records assigned to each cluster. Similarities were determined from calculating cosine similarity scores, with comparisons limited to the top 500 records per cluster. Higher similarity score represent greater overlap in text content between statistical methods sections. Results were transformed to lower case for the clustering, but examples are given using the original capitalisation.

@@

<!-- ```{r,fig.width=10,fig.cap="\\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters"} -->

<!-- include_graphics("figures/plosone.10topics.jpg") -->

<!-- ``` -->
